Automated topic naming to support analysis of software maintenance activities

### 2010-08-19 icse/icse2011.tex
moving comments on methodology up in the outline, closer to section it came from

\subsection{Generating the data}
\label{sec:wordlist}

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, so the functional requirements would be broadly similar. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains 320 thousand lines of C and C++ source code.  
Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. 
This is the most readily accessible source of project interactions for outside researchers; bug trackers and email-list archives were not available for both projects. 
An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. 
We extracted these messages and indexed them by creation time. 
We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}. 
Stemming was performed in the later stages of our analysis. % WAS IT?

For the commit message data-sets of each project, we created an XML file which separated commits into 30 day periods. 
This size of period, 30 days, is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. 
For each period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis algorithm~\cite{Blei2003}, and recorded the topics of each period.
%For each project and for each period of messages a project, we input these messages 
%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. 
Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. 
LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data, word distributions of messages, with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. 
Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
A topic, in this paper, represents both a word distribution and a group of commit log comments that are related to each other by their content.  
In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
%\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}

@@

Based on these results we find that \emph{reliability} and \emph{usability} worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}. MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as \emph{functionality} did improve. If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our results were still better than random chance.

@@

@@

For example, in Fig. \ref{fig:mysql-timeline}, we see that the NFRs functionality, portability and maintainability contain more labeled topics, since these NFRs have more intense shaded boxes.

@@

\subfloat[MaxDB 7.500]{                                 
\includegraphics[width=\textwidth]{figures/maxdb-timeline}
\label{fig:mysql-timeline}
}
\caption[]{NFR label per period. Each box represents a 30-day period. Numbers represent number of topics in that period labelled with that NFR. Box intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs. Grey bars indicate label frequency relative to \emph{that} NFR.
% \todo[inline]{better discussion}
}
\label{fig:timelines}
\end{figure*}

@@

\subsection{Annotation observations}

We found many topics that were not non-functional requirements (NFRs) but were often related to them. For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because concurrency was troublesome. Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management. These kinds of changes are slightly related to \emph{maintainability}. A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.

@@

% para 1 we did this, we were successful
In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling non-functional requirements topics and artifacts from project histories.
Our contributions include:
\begin{itemize}
\item We provide an unsupervised method of topic labelling (word-bags);
%\item Our unsupervised method can be applied to projects without training;
\item We demonstrate a supervised method of labelling topics by a single NFRs (machine learning);
\item We demonstrate a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
\item We provide a method of cross-project analysis via topic labelling;
\item We demonstrate a method of visualizing these topics across time.
\end{itemize}
We leveraged word-bags, machine learning classifiers and multi-label machine learning classifiers.  
We evaluated these techniques using multiple experiments against two open source database systems and visualized the results.
We found that our topic labelling techniques were effective.

>>>>

\subsection{Generating the data}
\label{sec:wordlist}
To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. We selected systems from the same application domain, so the functional requirements would be broadly similar. We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.  Choosing an older version of these projects allows us to focus on projects which have moved into the maintenance phase of the software life-cycle.

For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. This is the most readily accessible source of project interactions for outside researchers; bug reports and email messages were not available for both projects. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}. Stemming was performed in the later stages of our analysis.

From that data-set, we created an XML file which separated commits into 30 day periods. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm~\cite{Blei2003}.

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then used to recreate the word distributions of the underlying messages. These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).
\todo[inline]{These paragraphs should've come way earlier, in previous we should explain LDA so this doesn't need to be here.}

@@

Based on these results we find that \emph{reliability} and \emph{usability} worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}. MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as \emph{functionality} did improve. If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our classifier nonetheless performed substantially better than random.

@@

\subsection{Annotation observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, because concurrency was troublesome. Configuration management and source control related changes appeared often and sometimes there were topics dedicated to configuration management. These kinds of changes are slightly related to \emph{maintainability}. A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files.

@@

For example, in Fig. \ref{fig:mysql-timeline}, we see that the NFRs \emph{functionality}, \emph{portability} and \emph{maintainability} contain more labeled topics, since these NFRs have more intense shaded boxes.

@@

\subfloat[MaxDB 7.500]{                                 
\includegraphics[width=\textwidth]{figures/maxdb-timeline}
\label{fig:mysql-timeline}
}
\caption[]{NFR label per period. Each box represents a 30-day period. Numbers represent number of topics in that period labelled with that NFR. Box intensity (saturation) is mapped to label frequency relative to the largest label value of \emph{all} NFRs. Grey bars indicate label frequency relative to \emph{a particular} NFR.
% \todo[inline]{better discussion}
}
\label{fig:timelines}
\end{figure*}

@@

@@

% para 1 we did this, we were successful
In this paper we presented a cross-project data mining technique that was appropriate for mining and labelling latent software development topics using a non-functional requirements taxonomy. %topics and artifacts from project histories.
Our contributions include:
\begin{itemize}
\item We provide an unsupervised method of topic labelling (word-bags);
%\item Our unsupervised method can be applied to projects without training;
\item We demonstrate a supervised method of labelling topics by a single NFRs (machine learning);
\item We demonstrate a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
\item We provide a method of cross-project analysis via topic labelling;
\item We demonstrate a method of visualizing these topics across time.
\end{itemize}
We leveraged word-bags, machine learning classifiers and multi-label machine learning classifiers.  
We evaluated these techniques using multiple experiments against two open source database systems and visualized the results.
We found that our topic labelling techniques were effective.
